\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{A Radial Time Projection Chamber for $\alpha$ detection in CLAS at JLab}

### 2017-07-07 eg6-letters/NIM/EG6_RTPC_NIM.tex
NIM:  add section on eloss, capitalize helium in various places

For certain experiments the base CLAS system was complimented with ancillary 
detectors. For example, the measurement of the Deeply Virtual Compton 
Scattering (DVCS) process ($eH \rightarrow e' H' \gamma$, where $H$ is a 
nucleon or nuclei) necessitates an upgrade of the photon detection system. 
Indeed, with a 6 GeV electron beam, the majority of DVCS photons are produced 
at very forward angles, where the acceptance of the EC was poor. To extend the 
detection range, an inner calorimeter (IC) was built for the E01-113 
experiment in 2005~\cite{FX}. The IC was constructed from 424 lead-tungstate 
(PbWO$_{4}$) crystals, covering polar angles between 5$^{\circ}$ and 
15$^{\circ}$~\cite{Hyon-suk}. To protect the 
CLAS detector and the IC from the large flux of low energy M{\o}ller 
electrons, a 5~T solenoid magnet was placed around the target to shield the detectors. 
To detect recoiling $\alpha$ particles from the coherent DVCS on helium, a new 
radial time projection chamber (RTPC) was developed to track low energy nuclear 
fragments. The solenoid field was used to track and measure momentum of 
particles in the RTPC. The CLAS detector supplemented with both IC and RTPC was used 
during a three months experimental run~\cite{proposal1,proposal2}
in 2009 with a longitudinally polarized electron beam of 130~nA 
and energy of 6.064 GeV incident on a gaseous $^{4}$He target.

@@

We detail here the different regions shown in Fig.~\ref{fig:RTPC_1_4} starting 
from the beam line towards larger radius:\\
\begin{itemize}
   \item The 6 atm helium-4 target extends along the beam axis; it 
      is a 284~mm long, 6~mm diameter Kapton straw with a 27-$\mu$m wall.
   \item The first gas gap covers the radial range from 3 mm to 20 mm. It is 
      filled with $^{4}$He gas at 1~atm to minimize secondary interactions from
      M\o{}ller electrons scattered by the beam. This region is surrounded by 
      a 4 $\mu$m thick window made of grounded aluminized Mylar.
   \item The second gap region extends between 20~mm and 30~mm and is filled with the 
      gas mixture of 80$\%$ neon (Ne) and 20$\%$ dimethyl ether (DME). This region 
      is surrounded by a 4~$\mu$m thick window made of aluminized Mylar set at $- 4260$~V 
      to serve as the cathode.
   \item The drift region is filled with the same Ne-DME gas mixture and extends 
      from the cathode to the first GEM, 60 mm away from the beam axis. The average 
      electric field in this region is perpendicular to the beam and about 550~V/cm.
   \item The electron amplification system is composed of three GEMs located at 
      radii of 60, 63 and 66 mm. In this configuration, the first GEM layer 
      serves as the anode and each subsequent GEM is set at a lower voltage to
      obtain a strong ($\sim$1600 V/cm) electric field between the GEM foils. A 
      275~V bias is applied across each GEM for amplification.
   \item The readout board has an internal radius of 69~mm and collects charge
     from the GEMs. Pre-amplifiers are plugged directly on its outer side and
     transmit the signal to the data acquisition electronics.
\end{itemize}

@@

Based on the kinematics of the electrons in the calibration data, 
we generated the helium nucleus in our RTPC GEANT4 simulation~\cite{GEANT4}. Then,
we compared the calculated GEANT4 trajectory of the Helium nuclei to 
the hits measured in the chamber. To perform this drift path extraction, 
we made a first approximation assuming a linear dependence between the radius 
of emission of the charge and its time of detection, and then refined our 
result. Indeed, because of the magnetic field, the drift paths are not linear 
in the RTPC, but the curvature was minimal and this process converged already 
on the second iteration.

@@

@@

We reported on the construction, operation and calibration of a small RTPC 
designed to measure helium-4 nuclei in high rate environment. The operation
of the detector was successful and allowed to detect helium nuclei with a 75\%
efficiency and a readout rate
of 3.1 kHz triggered by the detection of high energy electrons and 
photons in the CLAS spectrometer.

>>>>

For certain experiments the base CLAS system was complimented with ancillary 
detectors. For example, the measurement of the Deeply Virtual Compton 
Scattering (DVCS) process ($eH \rightarrow e' H' \gamma$, where $H$ is a 
nucleon or nuclei) necessitates an upgrade of the photon detection system. 
Indeed, with a 6 GeV electron beam, the majority of DVCS photons are produced 
at very forward angles, where the acceptance of the EC was poor. To extend the 
detection range, an inner calorimeter (IC) was built for the E01-113 
experiment in 2005~\cite{FX}. The IC was constructed from 424 lead-tungstate 
(PbWO$_{4}$) crystals, covering polar angles between 5$^{\circ}$ and 
15$^{\circ}$~\cite{Hyon-suk}. To protect the 
CLAS detector and the IC from the large flux of low energy M{\o}ller 
electrons, a 5~T solenoid magnet was placed around the target to shield the detectors. 
To detect recoiling $\alpha$ particles from the coherent DVCS on Helium, a new 
radial time projection chamber (RTPC) was developed to track low energy nuclear 
fragments. The solenoid field was used to track and measure momentum of 
particles in the RTPC. The CLAS detector supplemented with both IC and RTPC was used 
during a three months experimental run~\cite{proposal1,proposal2}
in 2009 with a longitudinally polarized electron beam of 130~nA 
and energy of 6.064 GeV incident on a gaseous $^{4}$He target.

@@

We detail here the different regions shown in Fig.~\ref{fig:RTPC_1_4} starting 
from the beam line towards larger radius:\\
\begin{itemize}
  \item The 6-atm $^4$He target extends along the beam axis; it 
      is a 284-mm-long, 6-mm-diameter Kapton straw with a 27-$\mu$m wall.
   \item The first gas gap covers the radial range from 3 mm to 20 mm. It is 
      filled with $^{4}$He gas at 1~atm to minimize secondary interactions from
      M\o{}ller electrons scattered by the beam. This region is surrounded by 
      a 4 $\mu$m thick window made of grounded aluminized Mylar.
   \item The second gap region extends between 20~mm and 30~mm and is filled with the 
      gas mixture of 80$\%$ neon (Ne) and 20$\%$ dimethyl ether (DME). This region 
      is surrounded by a 4~$\mu$m thick window made of aluminized Mylar set at $- 4260$~V 
      to serve as the cathode.
   \item The drift region is filled with the same Ne-DME gas mixture and extends 
      from the cathode to the first GEM, 60 mm away from the beam axis. The average 
      electric field in this region is perpendicular to the beam and about 550~V/cm.
   \item The electron amplification system is composed of three GEMs located at 
      radii of 60, 63 and 66 mm. In this configuration, the first GEM layer 
      serves as the anode and each subsequent GEM is set at a lower voltage to
      obtain a strong ($\sim$1600 V/cm) electric field between the GEM foils. A 
      275~V bias is applied across each GEM for amplification.
   \item The readout board has an internal radius of 69~mm and collects charge
     from the GEMs. Pre-amplifiers are plugged directly on its outer side and
     transmit the signal to the data acquisition electronics.
\end{itemize}

@@

Based on the kinematics of the electrons in the calibration data, 
we generated the Helium nucleus in our RTPC GEANT4 simulation~\cite{GEANT4}. Then,
we compared the calculated GEANT4 trajectory of the Helium nuclei to 
the hits measured in the chamber. To perform this drift path extraction, 
we made a first approximation assuming a linear dependence between the radius 
of emission of the charge and its time of detection, and then refined our 
result. Indeed, because of the magnetic field, the drift paths are not linear 
in the RTPC, but the curvature was minimal and this process converged already 
on the second iteration.

@@

\subsection{Energy Loss Corrections}\label{sec_eloss}
Energy loss between the target and drift region was significant in the RTPC and necessitated a correction for optimal momentum reconstruction at the primary interaction vertex.  The dominant loss was in the 27-$\mu$m-thick Kapton target wall, with significant contributions also from the pressurized target gas and the foils before reaching the drift region.  Corrections were developed based on GEANT4 simulations with the full RTPC geometry and parameterized in terms of recoil curvature in the drift region and polar angle, separetely for all recoil hypotheses ($p$, $d$, $^3$H, $^3$He, $^4$He).  At our average coherent $^4$He DVCS kinematics, energy losses were about 5 MeV, while for $e-^4$He elastic scattering losses were about 3 MeV, which corresponds to momentum corrections of 25\% and 15\%, respectively.

@@

We reported on the construction, operation and calibration of a small RTPC 
designed to measure Helium-4 nuclei in high rate environment. The operation
of the detector was successful and allowed to detect Helium nuclei with a 75\%
efficiency and a readout rate
of 3.1 kHz triggered by the detection of high energy electrons and 
photons in the CLAS spectrometer.
